\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{cancel}
\usepackage{enumitem}
\usepackage{esdiff}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{ulem}
% \usepackage{pgfplots}
\usepackage{wrapfig}
\usepackage{xcolor}

\newcommand{\e}[1]{e^{#1}}
\newcommand{\ei}[1]{e^{i\left(#1\right)}}
\newcommand{\E}[1]{\times 10^{#1}}

\renewcommand\qedsymbol{TENA}

\title{
    Homework \#5
    \\  \small
    PHYS 4D: Modern Physics
    }
\author{Donald Aingworth IV}
% \date{January 26, 2026}

\begin{document}
    \maketitle

    \section{Questions}
    \begin{multicols}{2}
        \small
        \subsection{Question 1}
            Give the operators corresponding to the momentum and energy of a particle.

            \subsubsection{Solution}
                \begin{equation}
                    \hat{p} = -i \hbar \diffp{}{x};
                    \hat{H} = -\frac{\hbar^2}{2m} \diffp[2]{}{t} + V(x,t)
                \end{equation}
        
        \subsection{Question 2}
            Give the general form of an eigenvalue equation.

            \subsubsection{Solution}
                For operator $\hat{A}$, eigenfunction $\psi$, and eignenvalue $A$.
                \begin{equation}
                    \hat{A}\psi = A\psi
                \end{equation}

        \subsection{Question 3}
            What significance do the eigenvalues have?

            \subsubsection{Solution}
                The eigenvalue is the formula for or value of the component being searched for within the bounds of the eigenfunction.

        \subsection{Question 4}
            What significance do the eigenfunctions have?
                
            \subsubsection{Solution}
                The eigenfunctions serve as the conditions governing the system.

        \subsection{Question 5}
            Write down the momentum eigenvalue equation.
            
            \subsubsection{Solution}
                Ths is a \textit{solved} momentum eigenvalue equation.
                \begin{equation}
                    -i \hbar \diffp{}{x} \psi = \hbar k \psi
                \end{equation}

        \subsection{Question 6}
            Is the function $\cos(kx)$ an eigenfunction of the momentum operator?

            \subsubsection{Solution}
                No. 
                The first derivative of $\cos(kx)$ is $-k\sin(kx)$, which does not contain $\cos(kx)$ to the point that the sets of terms in the two equations are practically disjoint.

        \subsection{Question 7}
            Is the function $\cos(kx)$ an eigenfunction of the operator corresponding to the kinetic energy?

        \subsection{Question 8}
            An electron is described by the wave function $\phi(x)=Ae^{i \alpha x}$, where $\alpha$ denotes the Greek letter alpha. What is the momentum of the electron?

        \subsection{Question 9}
            How would you determine whether or not a particular wave function represented a state of the system having a definite value of the momentum?

        \subsection{Question 10}
            Calculate the dot product of the following two vectors.
            \begin{equation}
                A = \begin{pmatrix} 4 \\ 3 \\ 2 \\ 1 \end{pmatrix}; 
                B = \begin{pmatrix} 1 \\ 1 \\ 2 \\ 2 \end{pmatrix}
            \end{equation}

        \subsection{Question 11}
            For the two vectors, A and B, given in the preceding question, evaluate matrix multiplication $BA$ and from there $A \cdot BA$.

        \subsection{Question 12}
            Suppose a function f(x) is defined in an interval between 0 and 10 and an equally spaced grid is created by the command,\\
                {\fontfamily{qcr}\selectfont x = linspace(0.0,10.0,10)}
            and suppose the values of the function at the two Gauss quadrature points, $\xi_{i1}$ and $\xi_{i2}$, within each interval are known.
            Write down a formula for determining the value of the following integral in terms of the values of the function at the Gauss points.
            \begin{equation}
                \int_{0}^{10} f(x)\,dx
            \end{equation}

        \subsection{Question 13}
            Suppose a differential equation is solved using spline collocation on a grid created by the command {\fontfamily{qcr}\selectfont x = linspace(0.0,10.0,10)}.
            How many rows and columns do the A, B, and C matrices have?

        \subsection{Question 14}
            Sketch the wave function for an electron incident upon a potential step when the energy $E$ of the electron is greater than the step height $V_0$.

        \subsection{Question 15}
            Sketch the wave function for an electron incident upon a potential step when the energy $E$ of the electron is less than the step height $V_0$.

        \subsection{Question 16}
            Is the wavelength of a particle that has passed over a potential barrier greater than or less than the wavelength of the incident particle?

        \subsection{Question 17}
            Sketch the wave function of an electron which tunnels through a potential barrier located between 0 and L.

        \subsection{Question 18}
            Write down the Heisenberg uncertainty relations for the position and the momentum and for the time and the energy.

        \subsection{Question 19}
            Use the Heisenberg relation for the time and the energy to describe how the energy profile of an excited atomic state depends upon the lifetime of the state.

        \subsection{Question 20}
            How would the wave function $\psi(x)$ of a particle change as the width of the Fourier transform increases?

        \subsection{Question 21}
            Suppose that a particle localized between a and b is described by the wave function $\psi(x)$. Write down an equation for the average value of the kinetic energy of the particle.
            
    \end{multicols}
\end{document}