\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{cancel}
\usepackage{enumitem}
\usepackage{esdiff}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{ulem}
% \usepackage{pgfplots}
\usepackage{wrapfig}
\usepackage{xcolor}

\newcommand{\e}[1]{e^{#1}}
\newcommand{\ei}[1]{e^{i\left(#1\right)}}
\newcommand{\E}[1]{\times 10^{#1}}
\newcommand{\tagref}[1]{\tag{\ref{#1}}}

\renewcommand\qedsymbol{TENA}

\title{
    Homework \#10
    \\  \small
    PHYS 4D: Modern Physics
    \\
    Problems from Krane's textbook
    }
\author{Donald Aingworth IV}
\date{March 23, 2026}

\begin{document}
    \maketitle

    \section{Optional}
        \subsection{Chapter 7, Question 1}
            How does the quantum-mechanical interpretation of the hydrogen atom differ from the Bohr model?

            \subsubsection{Answer}
                The Quantum Mechanical model uses the particle-wave duality of quantum particles and treats electron positions in shells rather than in orbits.
                It also treats position in a probablistic manner rather than a definite manner.

        \subsection{Chapter 7, Question 3}
            What are the meanings of the quantum numbers $n, \ell, m_\ell$ according to (a) the quantum-mechanical calculation; (b) the vector model; (c) the Bohr (orbital) model?

            \subsubsection{Answer (a) and (b)}
                \begin{equation}
                    \psi = R_{n,\ell}(r) \Theta_{\ell, m_\ell}(\theta) \Phi_{m_\ell}(\phi)
                \end{equation}
                $n$, within the context of quantum mechanics, determines energy levels.
                It affects the radial position of the electron.
                The energy level is described by equation (\ref{eq:1}).
                \begin{equation}
                    \label{eq:1}
                    E_n = -\frac{me^4}{32\pi^2 \varepsilon_0^2 \hbar^2} \frac{1}{n^2}
                \end{equation}

                $\ell$ is used to calculate the magnitude of the angular momentum.
                It also affects the magnitude and planar position of the electron.
                The angular momentum magnitude is found using equation (\ref{eq:2}).
                \begin{equation}
                    \label{eq:2}
                    |\textbf{L}| = \sqrt{\ell(\ell + 1)}\hbar
                \end{equation}

                $m_\ell$ is used to calculate the magnitude of the z-component of the angular momentum.
                It also affects the planar angle and vertical position of the electron.
                The angular momentum z-component is found using equation (\ref{eq:3}).
                \begin{equation}
                    \label{eq:3}
                    L_z = m_\ell \hbar
                \end{equation}

            \subsubsection{Answer (c)}
                In the Bohr model, $n$ represents both the distance from the center (as in (\ref{eq:1})) and the angular momentum of the electron ($L = n\hbar$).
                It also determines energy levels.
                $\ell$ and $m_\ell$ represent nothing.
            
        \subsection{Chapter 7, Question 5}
            How does the orbital angular momentum differ between the Bohr model and the quantum-mechanical calculation?

            \subsubsection{Answer}
                In Bohr, it is a function of only the principal quantum number, so $L = n\hbar$.
                In QM, we only know the magnitude and the z-component of $L$.
                They are described by (\ref{eq:2}) and (\ref{eq:3}).
                \begin{gather}
                    \tag{\ref{eq:2}}
                    |\textbf{L}| = \sqrt{\ell(\ell + 1)}\hbar\\
                    \tag{\ref{eq:3}}
                    L_z = m_\ell \hbar
                \end{gather}
            
        \subsection{Chapter 7, Question 11}
            The probability density $\psi^* \psi$ does not depend on $\phi$ for the wave functions listed in Table 7.1. 
            What is the signiﬁcance of this?

            \subsubsection{Answer}
                Its significance is that the relevant values cancel out each other.
                Notably, the Heisenberg Uncertainty Principle says we can't know both position and momentum beyond a certain level.
                Since we already know the z-value of angular momentum to an exact degree (see equation (\ref{eq:3})), it makes sense that we would have no idea what the one value that affects z-position is.
            
        \subsection{Chapter 7, Question 13}
            Can a hydrogen atom in its ground state absorb a photon (of the proper energy) and end up in the 3d state?

            \subsubsection{Answer}
                No. 
                The rules are that $\Delta \ell = \pm 1$ and $\Delta m_\ell \in \{-1, 0, 1\}$.
                This involves transitioning for $\Delta \ell = 2$, which is more than the allowed value(s).
            
        \subsection{Chapter 7, Question 15}
            The photon has a spin quantum number of 1, but its spin magnetic moment is zero. 
            Explain.

            \subsubsection{Answer}
                For a thing to have a spin magnetic moment, it must have a charge.
                A photon does not have a charge, and also does not have mass.
                Therefore, it has no spin magnetic moment.
            
        \subsection{Chapter 7, Question 19}
            How would (a) the Zeeman effect and (b) the fine structure be different in a muonic hydrogen atom? 
            (See Problems 41 and 42 in Chapter 6.) 
            The muon has the same spin as the electron, but is 207 times as massive.

            \subsubsection{Answer (a)}
                Recall the formula for the orbital magnetic moment.
                \begin{equation}
                    \label{eq:1.7.1}
                    \vec{\mu}_L = -\frac{e}{2m}\vec{L}
                \end{equation}

                The orbital magnetic moment is inversely dependant on the mass of the particle.
                Since the muon mass is 207 times the electrom mass, the magnetic moment is 207 times smaller.
                The change in total energy is described in (\ref{eq:1.7.2}).
                \begin{equation}
                    \label{eq:1.7.2}
                    U = -\vec{\mu}_L \cdot \vec{B} = \frac{e}{2m} \vec{L} \cdot \vec{B}
                \end{equation}

                From this, the change in energy with regard to the spin magnetic moment would be less.
                This means that the differences in the energy levels would be less.

            \subsubsection{Answer (b)}
                Recall the formula for the spin magnetic moment.
                \begin{equation}
                    \label{eq:1.7.3}
                    \vec{\mu}_S = -\frac{e}{m}\vec{S}
                \end{equation}

                The spin magnetic moment is inversely dependant on the mass of the particle.
                Since the muon mass is 207 times the electrom mass, the magnetic moment is 207 times smaller.
                The change in total energy is described in (\ref{eq:1.7.2}).
                \begin{equation}
                    \label{eq:1.7.4}
                    U = -\vec{\mu}_L \cdot \vec{B} = \frac{e}{m} \vec{S} \cdot \vec{B}
                \end{equation}

                From this, the change in energy with regard to the spin magnetic moment would be less.
                This means that the levels of the splitting of the energy levels would be less.
            \pagebreak

        \subsection{Chapter 7, Problem 21}
            List the excited states (in spectroscopic notation) to which the $4p$ state can make downward transitions.

            \subsubsection{Solution}
                We have a rule for this.
                \begin{equation}
                    \label{eq:1.8.1}
                    \Delta \ell = \pm 1
                \end{equation}

                \noindent
                We start from $4p$, or $n = 4, \ell = 1$.
                It can go downward to $1s$, $2s$, $3s$, or $3d$.
            
        \subsection{Chapter 7, Problem 22}
            (a) A hydrogen atom is in an excited $5g$ state, from which it makes a series of transitions by emitting photons, ending in the $1s$ state. 
            Show, on a diagram similar to Figure 7.19, the sequence of transitions that can occur. 
            (b) Repeat part (a) if the atom begins in the $5d$ state.

            \subsubsection{Solution}
                I'll just list it. 
                No drawing.
                \begin{align*}
                    (a):\,& 5g \to 4f \to 3d \to 2p \to 1s\\
                    (b):\,& 5d \to 3p \to 1s
                \end{align*}
            \pagebreak

        \subsection{Chapter 7, Problem 25}
            A collection of hydrogen atoms is placed in a magnetic field of 3.50 T. 
            Ignoring the effects of electron spin, find the wavelengths of the three normal Zeeman components (a) of the 3d to 2p transition; (b) of the 3s to 2p transition.

            \subsubsection{Solution (a)}
                To answer this, we must find the change in energy.
                \begin{equation}
                    \Delta E = E_f - E_i
                \end{equation}

                The base energy comes from equation (\ref{eq:1}).
                \begin{equation}
                    \tag{\ref{eq:1}}
                    E_n = -\frac{me^4}{32\pi^2 \varepsilon_0^2 \hbar^2} \frac{1}{n^2}
                        \approx -13.6\,\unit{\eV}\,\frac{Z^2}{n^2}
                \end{equation}

                We can apply this for Hydrogen ($Z = 1$) to this specific case.
                \begin{align}
                    \Delta E    &=  E_f - E_i
                        =   (-13.6\,\unit{\eV}) Z^2 \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right)\\
                        &=  (-13.6\,\unit{\eV}) \left( \frac{1}{2^2} - \frac{1}{3^2} \right)
                        =   (-13.6\,\unit{\eV}) * \frac{5}{36}
                        =   -\frac{17}{9}\,\unit{\eV}
                \end{align}

                This value of energy is usable for parts (a) and (b).
                We can add that on in the cases of the change in energy from the magnetic field.
                They come from potential energy and can be calculated from equation (\ref{eq:7.25.4}).
                \begin{equation}
                    \label{eq:7.25.4}
                    \Delta U = \Delta m_\ell \mu_B B
                \end{equation}

                There are only three possible values of $\Delta m_\ell$.
                \begin{equation}
                    \Delta m_\ell \in \{ -1, 0, +1 \}
                \end{equation}

                This can be applied in three cases.\\
                $\Delta m_\ell = 1$:
                \begin{align}
                    E   &=  \Delta E + U
                        =   (-13.6\,\unit{\eV}) Z^2 \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right) + \Delta m_\ell \mu_B B\\
                        &=  \frac{17}{9}\,\unit{\eV} * \frac{1.602\E{-19}\,\unit{\joule}}{1\,\unit{\eV}} + 1 * (9.274\E{-24}\,\unit{\joule/\tesla}) (3.50\,\unit{\tesla})\\
                        &=  302.6\E{-21}\,\unit{\joule} + 32.459\E{-24}\,\unit{\joule}
                        =   302.632\E{-21}\,\unit{\joule}\\
                    \lambda &=  \frac{hc}{E}
                        =   \frac{hc}{302.632\E{-21}\,\unit{\joule}}
                        =   \boxed{656.40\,\unit{\nano\meter}}
                \end{align}

                $\Delta m_\ell = 0$:
                \begin{align}
                    E   &=  \Delta E + U
                        =   (-13.6\,\unit{\eV}) Z^2 \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right) + \Delta m_\ell \mu_B B\\
                        &=  \frac{17}{9}\,\unit{\eV} * \frac{1.602\E{-19}\,\unit{\joule}}{1\,\unit{\eV}} + 0 * (9.274\E{-24}\,\unit{\joule/\tesla}) (3.50\,\unit{\tesla})\\
                        &=  302.6\E{-21}\,\unit{\joule} + 0
                        =   302.6\E{-21}\,\unit{\joule}\\
                    \lambda &=  \frac{hc}{E}
                        =   \frac{hc}{302.6\E{-21}\,\unit{\joule}}
                        =   \boxed{656.47\,\unit{\nano\meter}}
                \end{align}

                $\Delta m_\ell = -1$:
                \begin{align}
                    E   &=  \Delta E + U
                        =   (-13.6\,\unit{\eV}) Z^2 \left( \frac{1}{n_f^2} - \frac{1}{n_i^2} \right) + \Delta m_\ell \mu_B B\\
                        &=  \frac{17}{9}\,\unit{\eV} * \frac{1.602\E{-19}\,\unit{\joule}}{1\,\unit{\eV}} - 1 * (9.274\E{-24}\,\unit{\joule/\tesla}) (3.50\,\unit{\tesla})\\
                        &=  302.6\E{-21}\,\unit{\joule} - 32.459\E{-24}\,\unit{\joule}
                        =   302.632\E{-21}\,\unit{\joule}\\
                    \lambda &=  \frac{hc}{E}
                        =   \frac{hc}{302.632\E{-21}\,\unit{\joule}}
                        =   \boxed{656.54\,\unit{\nano\meter}}
                \end{align}
            \pagebreak

    %     \subsection{Chapter 7, Problem 27}
    %         Calculate the energies and wavelengths of the 3d to 2p transition, taking into account the ﬁne structure of both levels.
    %         How many component wavelengths might there be in the transition?

    %         \subsubsection{Solution}
    %             Discounting spin, there were only three.
    %             Counting spin, that duplicates energy values and moves them up or down by $\mu_B B$.

            
    %     \subsection{Chapter 7, Problem 34}
    %         For the 1s, 2s, and 2p states of hydrogen, show that $(r^{-1})_{\rm av} = 1/n^2a_0$. 
    %         This turns out to be a general result for any state of hydrogen. 
    %         Based on this result, explain why the Bohr model gives such a good estimate for the ﬁne-structure splitting as well as for other magnetic effects due to the circulating electron.

    %         \subsubsection{Solution}
            
    % \pagebreak
    
    \section{Chapter 8, Question 2}
        Why do the $4s$ and $3d$ subshells appear so close in energy, when they belong to different principal quantum numbers $n$?

        \subsection{Answer}
            This is because the the large number of electrons in the inner shells in the $n = 3$ shell cause a lot of screening.
            Both the $4s$ and $3d$ are affected by this screening, to the point that they have around the same energy.
            They are not exactly the same energy, and the amount of energy of each depends on the element, which is what results in the $3d$ shell of Copper filling up faster than its $4s$ shell. 
    % \pagebreak
    
    \section{Chapter 8, Question 3}
        Would you expect element 107 to be a good conductor or a poor conductor? 
        How about element 111? 
        Do you expect element 112 to be paramagnetic or diamagnetic?

        \subsection{Answer}
            Element 107 (Bh) has a configuration of $\rm [Rn] 7s^2 5f^{14} 6d^5$.
            Transition metals with half-filled d-shells have higher resistivities than their neighbors.
            Therefore, I expect it to have a high resistivity, but not an absurdly high one. 
            This means a \underline{poor conductor}.

            Element 111 (Rg) has a configuration of $\rm [Rn] 7s^2 5f^{14} 6d^{9}$.
            I expect it to be a much better conductivity than Bohrium.
            This means a \underline{good conductor}.

            Element 112 has a configuration of $\rm [Rn] 7s^2 5f^{14} 6d^{10}$.
            It has only full shells.
            This means \underline{diamagnetic}.
    \pagebreak

    \section{Chapter 8, Question 10}
        Suppose we do a Stern-Gerlach experiment using an atom that has angular momentum quantum numbers $L$ and $S$ in its ground state. 
        Into how many components will the beam split? 
        Do you expect them to be equally spaced?

        \subsection{Answer}
            Having to do with magnetic field coming from motion and spin of electrons, motion from a magnet would come from different values of $L$ and $S$.
            There are two kinds of spin, one that comes from the angular momentum ($L$) and one that comes from the spin ($S$). 
            There is a total rotational, which we can denote $J$ and which would be the sum of the angular momentum and the spin.
            \begin{equation}
                \vec{J} = \vec{L} + \vec{S}
            \end{equation}

            From $L$, there are $2L + 1$ possible values of $m_\ell$.
            From $S$, there are two possible values of $m_s$ for electrons, which are $\pm \frac{1}{2}$.
            If we add them together as vectors and take the absolute value, we get $J$.
            This results in a total of $2J + 1$ components, which we would expect to be equally spaced.
    \pagebreak
    
    \section{Chapter 8, Question 17}
        The first excited state in sodium is a ﬁne-structure doublet; the wavelengths emitted in the decay of these states are 589.59 nm and 589.00 nm, a difference of 0.59 nm.
        The excited $4s^1$ state in sodium (see Figure 8.4) decays to the $3p$ doublet with the emission of radiation at the wavelengths 1138.15 nm and 1140.38 nm, a difference of 2.23 nm. 
        Explain how the $3p$ ﬁne structure can give a wavelength difference of 0.59 nm in one case and 2.23 nm in the other case.

        \subsection{Solution}
            There are two parts to this, one rooted in Physics and one rooted in math.
            For Physics, we consider the idea of splitting energy levels for each electron.
            This splitting can result in a difference in the change in the energy that emits a photon.

            Even if the differences in the changes in energy were the same, the wavelengths would still be different. 
            Wavelength is inversely proportional to energy of the photon.
            Inverse proportionality has a whose different scale in change in proportionality.
            For example, while the difference between 3 and 4 is equal to the difference between 4 and 5, the difference between $\frac{1}{3}$ and $\frac{1}{4}$ is more than the difference between $\frac{1}{4}$ and $\frac{1}{5}$.
            This acts on the changes in energy and wavelength as well, resulting in larger wavelength spreads even if the energy spreads' sizes are identical.
    \pagebreak
    
    \section{Chapter 8, Question 20}
        Explain what is meant by a population inversion and why it is necessary for the operation of a laser.

        \subsection{Solution}
            A population inversion is when the proportions of energy levels swaps from mostly low and a few high to mostly high and a few low. 
            Atoms in the ground state can absorb the photons emitted by other atoms.
            If they are absorbed, they cannot exit the enclosure of the laser. 
            Without the population inversion, most of the photons released will be reabsorbed by other atoms.
            Having the population inversion lowers the possibility of the photons being reabsorbed and strengthens the amplitude of the laser.
    \pagebreak
    
    \section{Chapter 8, Problem 1}
        (a) List the six possible sets of quantum numbers $(n, \ell, m_\ell, m_s)$ of a $2p$ electron. 
        (b) Suppose we have an atom such as carbon, which has two $2p$ electrons. 
        Ignoring the Pauli principle, how many different possible combinations of quantum numbers of the two electrons are there? 
        (c) How many of the possible combinations of part (b) are eliminated by applying the Pauli principle? 
        (d) Suppose carbon is in an excited state with conﬁguration $2p^1 3p^1$. 
        Does the Pauli principle restrict the choice of quantum numbers for the electrons? 
        How many different sets of quantum numbers are possible for the two electrons?
        
        \subsection{Solution (a)}
            In every case, $n = 2$ and $\ell = 1$.
            The possible values of $m_\ell$ are $\{-1, 0, 1\}$ and the possible values of $m_s$ are $\{-\frac{1}{2}, \frac{1}{2}\}$.
            This gives us our six possible values.
            \begin{align*}
                &(2, 1, +1, +\frac{1}{2})   &(2, 1, +1, -\frac{1}{2})\\
                &(2, 1, 0, +\frac{1}{2})    &(2, 1, 0, -\frac{1}{2})\\
                &(2, 1, -1, +\frac{1}{2})   &(2, 1, -1, -\frac{1}{2})
            \end{align*}

        \subsection{Solution (b)}
            Use triangle numbers, to acount for repeats.
            In this case, it's the sixth triangle number (6 possible values).
            \begin{equation}
                C = \sum_{i = 1}^{6} i = 6+5+4+3+2+1 = \boxed{21}
            \end{equation}
        
        \subsection{Solution (c)}
            This is an $n$ choose $r$ problem where the order does matter. 
            \begin{equation}
                \binom{n}{r} = \frac{n!}{r!(n - r)!}
            \end{equation}

            In this case, $n = 6$ possible configurations and $r = 2$ electrons.
            \begin{align}
                C   &=  \binom{6}{2} 
                    =   \frac{6!}{2!(6 - 2)!}
                    =   \frac{6 * 5}{2}
                    =   15
            \end{align}

            This means that \boxed{6} combinations are eliminated.

        \subsection{Solution (d)}
            Pauli does restrict every electron, but in this case, the value of $n$ will be different for each electron, so there will be no other restrictions on values.
            One electron has $n = 2$, while the other has $n = 3$.
            $m_\ell$ has the same 6 possible values. 
            Due to the different values of $n$, there will be no repeats. 
            Multiple the number of possible values of each together.
            \begin{equation}
                C = 6 * 6 = \boxed{36}
            \end{equation}
    \pagebreak
    
    \section{Chapter 8, Problem 19}
        A small helium-neon laser produces a light beam with an average power of 3.5 mW and a diameter of 2.4 mm.
        (a) How many photons per second are emitted by the laser?
        (b) What is the amplitude of the electric ﬁeld of the light wave? Compare this result with the electric ﬁeld at a distance of 1 m from an incandescent light bulb that emits 100 W of visible light.

        \subsection{Solution (a)}
            The power can be converted to energy by multiplying by one second.
            \begin{equation}
                E_{\rm laser}   =   P * t 
                    =   3.5\,\unit{\milli\watt} * 1 \unit{\second}
                    =   3.5\,\unit{\milli\joule}
                    =   3.5\E{-3}\,\unit{\joule}
            \end{equation}

            A helium-neon laser has laser beams that come from the neon. 
            Said neon makes the transition from $2p^5 5s^1$ to $2p^5 3p^1$.
            The transition of an electron from $n = 5$ to $n = 3$ emits a photon, whose energy we can calculate.
            \begin{align}
                E_{\rm photon}  &=  -13.6\,\unit{\eV} * 10^2 * \left( \frac{1}{5^2} - \frac{1}{3^2} \right)
                    =   13.6\,\unit{\eV} * 10^2 * \frac{16}{225}\\
                    &=  13.6\,\unit{\eV} * \frac{64}{9} * \frac{1\,\unit{\joule}}{6.242\E{18}\,\unit{\eV}}
                    =   \frac{13.6\,\unit{\joule} * 64}{9 * 6.242\E{18}\,\unit{\eV}}\\
                    &=  1.549\E{-17}\,\unit{\joule}
            \end{align}

            If we divide the total energy by the energy of each photon, we will get the total number of photons, approximately. 
            \begin{align}
                n   &=  \frac{E_{\rm laser}}{E_{\rm photon}}
                    =   \frac{3.5\E{-3}\,\unit{\joule}}{1.549\E{-17}\,\unit{\joule}}
                    =   \boxed{2.26\E{14}}
            \end{align}

        \subsection{Solution (b)}
            The intensity light wave can be found with its power and its area concentrated to.
            \begin{equation}
                I = \frac{P}{A} = \frac{P}{\pi r^2} = \frac{3.5\,\unit{\milli\watt}}{\pi (1.2\,\unit{\milli\meter})^2}
            \end{equation}

            The electric field can be calcualted from the intensity.
            \begin{align}
                |\vec{E}|   &=  \sqrt{2I\mu_0 c}
                    =   \sqrt{\frac{2P\mu_0 c}{\pi r^2}}
                    =   \sqrt{\frac{8\E{-7}\,\unit{\newton/\ampere^2} * P c}{r^2}}\\
                    &=  \sqrt{\frac{8\E{-7}\,\unit{\newton/\ampere^2} * 3.5\,\unit{\milli\watt} * 2.998\E{8}\,\unit{\meter/\second}}{(1.2\,\unit{\milli\meter})^2}}\\
                    &=  \sqrt{\frac{80 * 3.5\E{-3} * 2.998}{1.44\E{-6}}\,\unit{\frac{\newton^2}{\coulomb^2}}}
                    =   \sqrt{\frac{5246500}{9}}\,\unit{\frac{\newton}{\coulomb}}
                    =   \boxed{763.5\,\unit{\newton/\coulomb}}
            \end{align}

            Meanwhile, intensity from an incandescent light is calculatable from the total power of the bulb and the distance from it.
            \begin{equation}
                I = \frac{P}{4\pi r^2}
            \end{equation}

            Plug this into the equation of the electric field and solve for it.
            \begin{align}
                |\vec{E}|   &=  \sqrt{2I\mu_0 c}
                    =   \sqrt{\frac{2 P \mu_0 c}{4\pi r^2}}
                    =   \sqrt{\frac{P \mu_0 c}{2 \pi r^2}}\\
                    &=  \sqrt{\frac{100\,\unit{\watt} * 4\pi\E{-7}\,\unit{\newton/\ampere^2} * 2.998\E{8}\,\unit{\meter/\second}}{2 \pi (1\,\unit{\meter})^2}}\\
                    &=  \sqrt{100\,\unit{\watt/\meter^2} * 20 * 2.998\,\unit{\newton\meter/\ampere^2\second}}\\
                    &=  \sqrt{2000 * 2.998\,\unit{\newton^2/\coulomb^2}}
                    =   \sqrt{5996}\,\unit{\newton/\coulomb}
                    =   77.43\,\unit{\newton/\coulomb}
            \end{align}

            One meter away, the electric field is about a tenth of the electric field of the laser.
    \pagebreak
    
    \section{Chapter 8, Problem 21}
        (a) The ionization energy of sodium is 5.14 eV. 
        What is the effective charge seen by the outer electron? 
        (b) If the $3s$ electron of a sodium atom is moved to the $4f$ state, the measured binding energy is 0.85 eV. 
        What is the effective charge seen by an electron in this state?

        \subsection{Solution (a)}
            On the periodic table, Sodium has the atomic number 11.
            Its ground state electron configuration is $\rm 1s^2 2s^2 2p^6 3s^1$. 
            This means that there is only one outer electron, for which $n = 3$.
            Use the formula for the energy of a screened electron to find the value of $Z_{\rm eff}$.
            Inverse the sign of the $13.6\,\unit{\eV}$ so we don't end up with an imaginary number.
            \begin{gather}
                E_n = (13.6\,\unit{\eV}) \frac{Z_{\rm eff}^2}{n^2} \to Z_{\rm eff}^2 = \frac{E_n n^2}{13.6\,\unit{\eV}}\\
                \begin{align}
                    Z_{\rm eff} &=  \sqrt{\frac{E_n n^2}{13.6\,\unit{\eV}}}
                        =   3\sqrt{\frac{5.14}{13.6}}
                        =   3\sqrt{0.378}
                        =   3 * 0.615
                        =   \boxed{1.84}
                \end{align}
            \end{gather}

        \subsection{Solution (b)}
            First subtract the binding energy from the ionization energy to find the ionization energy from the $4f$ state.
            \begin{equation}
                E_4 =   E_3 - \Delta E
                    =   5.14\,\unit{\eV} - 0.85\,\unit{\eV}
                    =   4.29\,\unit{\eV}
            \end{equation}

            From here, use the same equation as in part (a).
            \begin{align}
                Z_{\rm eff} &=  \sqrt{\frac{E_n n^2}{13.6\,\unit{\eV}}}
                    =   4\sqrt{\frac{4.29}{13.6}}
                    =   4\sqrt{3.15}
                    =   4 * 0.562
                    =   \boxed{2.25}
            \end{align}
\end{document}