\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cancel}
\usepackage{graphicx}
% \usepackage{physics}
\usepackage{siunitx}
\usepackage{wrapfig}

% \AtBeginDocument{\RenewCommandCopy\qty\SI}

\newcommand{\E}[1]{\times 10^{#1}}

\title{
    Chapter 17 End-of-Chapter Problems
    \\ \small
    Halliday \& Resnick, 10th Edition
}

\author{Donald Aingworth IV}

\date{\small Hit me where it Matters}

\begin{document}
    \DeclareSIUnit{\atm}{atm}
    \DeclareSIUnit{\cal}{\ cal}
    \DeclareSIUnit{\Cal}{\ Cal}
    \DeclareSIUnit{\calorie}{\ cal}
    \DeclareSIUnit{\Calorie}{\ Cal}
    \DeclareSIUnit{\celsiusdegree}{C^\circ}
    \DeclareSIUnit{\fahrenheit}{^\circ F}
    \DeclareSIUnit{\fahrenheitdegree}{F^\circ}
    \DeclareSIUnit{\torr}{\ torr}

    \maketitle

    \begin{center}
        \textit{Where needed in the problems, use}\\
        speed of sound in air = 343 m/s\\
        \textit{and}\\
        density of air = 1.21 \unit{\kilo\gram/\meter^3}\\
        \textit{unless otherwise specified.}
    \end{center}

    \pagebreak
    \section{Problem 1}
        Two spectators at a soccer game see, and a moment later hear, the ball being kicked on the playing field. 
        The time delay for spectator A is 0.23 s, and for spectator B it is 0.12 s. 
        Sight lines from the two spectators to the player kicking the ball meet at an angle of 90\unit{\degree}. 
        How far are (a) spectator A and (b) spectator B from the player?
        (c) How far are the spectators from each other?

        \subsection{Solution (a)}
            This is a simple question to answer.
            The distance traveled to A would be equal to the speed of sound times the time taken to travel the distance.
            \begin{align}
                x   &=  vt
                    =   (343\,\unit{\meter/\second})(0.23\,\unit{\second})
                    =   \boxed{78.89\,\unit{\meter}}
            \end{align}

    \pagebreak
    \section{Problem 3}
        When the door of the Chapel of the Mausoleum in Hamilton, Scotland, is slammed shut, the last echo heard by someone standing just inside the door reportedly comes 15 s later. 
        (a) If that echo were due to a single reflection off a wall opposite the door, how far from the door is the wall? 
        (b) If, instead, the wall is 25.7 m away, how many reflections (back and forth) occur?

        \subsection{Solution}

    \pagebreak
    \section{Problem 5}
        Earthquakes generate sound waves inside Earth.
        Unlike a gas, Earth can experience both transverse (S) and longitudinal (P) sound waves. 
        Typically, the speed of S waves is about 4.5 km/s, and that of P waves 8.0 km/s. 
        A seismograph records P and S waves from an earthquake. 
        The first P waves arrive 3.0 min before the first S waves. 
        If the waves travel in a straight line, how far away did the earthquake occur?

        \subsection{Solution}

    \pagebreak
    \section{Problem 7}
        A stone is dropped into a well. 
        The splash is heard 3.00 s later. 
        What is the depth of the well?

        \subsection{Solution}

    \pagebreak
    \section{Problem 11}
        Diagnostic ultrasound of frequency 4.50 MHz is used to examine tumors in soft tissue. 
        (a) What is the wavelength in air of such a sound wave? 
        (b) If the speed of sound in tissue is 1500 m/s, what is the wavelength of this wave in tissue?

        \subsection{Solution}

    \pagebreak
    \section{Problem 15}

        \subsection{Solution}

    \pagebreak
    \section{Problem 17}

        \subsection{Solution}

    \pagebreak
    \section{Problem 19}

        \subsection{Solution}

    \pagebreak
    \section{Problem 20}

        \subsection{Solution}

    \pagebreak
    \section{Problem 25}

        \subsection{Solution}

    \pagebreak
    \section{Problem 27}

        \subsection{Solution}

    \pagebreak
    \section{Problem 29}

        \subsection{Solution}

    \pagebreak
    \section{Problem 35}

        \subsection{Solution}

    \pagebreak
    \section{Problem 39}

        \subsection{Solution}

    \pagebreak
    \section{Problem 41}

        \subsection{Solution}

    \pagebreak
    \section{Problem 47}

        \subsection{Solution}

    \pagebreak
    \section{Problem 49}

        \subsection{Solution}

    \pagebreak
    \section{Problem 51}

        \subsection{Solution}

    \pagebreak
    \section{Problem 53}

        \subsection{Solution}

    \pagebreak
    \section{Problem 55}

        \subsection{Solution}

    \pagebreak
    \section{Problem 57}

        \subsection{Solution}

    \pagebreak
    \section{Problem 61}

        \subsection{Solution}

    \pagebreak
    \section{Problem 63}

        \subsection{Solution}

    \pagebreak
    \section{Problem 71}

        \subsection{Solution}

    \pagebreak
    \section{Problem 81}

        \subsection{Solution}

    \pagebreak
    \section{Problem 87}

        \subsection{Solution}

    \pagebreak
    \section{Problem 99}

        \subsection{Solution}

    \pagebreak
    \section{Problem 107}

        \subsection{Solution}

    \pagebreak

    \tableofcontents
\end{document}